---
title: "Data analysis"
author: "Luis Fernando Delgado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
library(asreml)
source("https://raw.githubusercontent.com/Cassava2050/PPD/main/utilities_tidy.R") # 

trial_interest <- "DMEPR"
year_interest <- 2023
```

## master_data to save the results

```{r}
master_data <- list()
```

## Load the tidy data

```{r}
trial_set_number = 1
# all files in the folder
list_file = list.files(here::here("output"))
# tidy data of the trials interested
sel_file = list_file[str_detect(list_file, "_tidy_data4analysis_") &
                       str_detect(list_file,
                                  paste(year_interest, trial_interest, sep=""))]
# the data we will use
sel_file_use = sel_file[1]

sel_file_use
trial1_tidy = read.csv(here::here("output", sel_file_use), header=TRUE,
                        stringsAsFactors = FALSE,
                        as.is=T,
                        check.names = FALSE)
if(trial_set_number == 1){
  trial_tidy_all = trial1_tidy
}

colnames(trial_tidy_all)
```

# Fixing synonimum accesion_names

```{r}
trial_tidy_all <- trial_tidy_all %>% 
  mutate(accession_name = recode_factor(accession_name,
                                 `CR52` = "CR52A4",        
                                `IITA-TMS-IBA-980581` = "IBA-980581",
                                `IBA920057` = "IBA-920057",
                               `CR13` = "CR13-8",
                               `R5` = "Rayong5",
                               `TMEB419` = "TMEB-419",
                               `Hanoi1` = "TMEB-419",
                               `Hanoi2` = "IBA-920057",
                               `Hanoi3` = "IBA-972205",
                               `Hanoi4` = "IBA-980505",
                               `Hanoi5` = "IBA-980581",
                               `IBA972205` = "IBA-972205",
                               `IITA-TMS-IBA980581` = "IBA-980581", 
                               `KU50-15` = "KU50_is_KM94", 
                               `KU50-H4` = "KU50_is_KM94",
                               `IBA980505` = "IBA-980505"))

# save fixed data
write.csv(trial_tidy_all, here::here("output", paste("01_", year_interest, trial_interest,
  "_tidy_data4analysis_", Sys.Date(), ".csv", sep = "")), row.names = FALSE)
```


## Obtain all the trait information using a cloud file (gitHub) -------

```{r}
trait_all <-
  read.csv("https://raw.githubusercontent.com/lfdelgadom/standar_col_names_CB/main/standar_col_names.csv") %>%
  select(analysis_col_name) %>%
  filter(str_detect(analysis_col_name, "obs_"))
trait_all_adj <- gsub("obs_", "", trait_all$analysis_col_name)
trait_all_adj = c(trait_all_adj,
                  "harvest_number_plan", "germination_perc",
                  "starch_yield_ha", "starch_content",
                  "CMD_incidence", "CMD_infect_plants")
trait_all_adj <- gsub("-", "_", trait_all_adj)
```

## Meta info.

```{r}
meta_all <-
  read.csv("https://raw.githubusercontent.com/lfdelgadom/standar_col_names_CB/main/standar_col_names.csv") %>%
  select(analysis_col_name) %>%
  filter(str_detect(analysis_col_name, "use_"))
meta_all_adj <- gsub("use_", "", meta_all$analysis_col_name)
meta_all_adj <- c(
  meta_all_adj,
  "check_released", "latitude", "longitude",
  "altitude", "department", "country",
  "ag_zone", "location_short"
)
```

## Select the observations for analysis

```{r}
names(trial_tidy_all) <- gsub("-", "_", names(trial_tidy_all))
analysis_trait <- names(trial_tidy_all)[names(trial_tidy_all) %in% trait_all_adj]
print("All the traits investigated:")
print(analysis_trait)
```

## Select the meta information for analysis

```{r}
meta_col <- names(trial_tidy_all)[names(trial_tidy_all) %in% meta_all_adj]
print("All the meta information:")
print(meta_col)
```

## Check the SD of each trait

```{r}
trial_rm_sd <- remove_no_var_tidy(my_dat = trial_tidy_all,
                   analysis_trait = analysis_trait,
                   meta_info = meta_col)
master_data[["mean_of_sd"]] = sd_mean
```

## Trait ideal

```{r}
no_traits_for_analysis <- c("stake_plant" , "planted_number_plot", 
                            "harvest_number", "root_weight_air", 
                            "root_weight_water", "harvest_number_plan",
                            "yield_ha_v2", "root_rot_perc", "shoot_weight_plot",
                            "root_rot_number"
)

no_variation_traits <- c() # "CAD_5mon", "CAD_7mon", "CAD_3mon", "lodging1_3_6mon"

no_traits_for_analysis <- c(no_variation_traits, no_traits_for_analysis)

trait_ideal <- analysis_trait[!analysis_trait %in% no_traits_for_analysis]
print("the trait ideal is:"); trait_ideal

trait_ideal %>% as.data.frame() %>% write.table("clipboard", sep = "\t", col.names = T, row.names = F)
```

# Genotypic correlation (Phenotypic values)

```{r}
correlation <- gg_cor(
  colours = c("red", "white", "blue"),
  data = trial_rm_sd[, trait_ideal],
  label_size = 4
  )

ggsave(paste("images\\pheno_corr", trial_interest, Sys.Date(), ".png", sep = "_"),
     plot = correlation, units = "in", dpi = 300, width = 12, height = 8
    )
```

## Check design experimental

### Agriutilities library

```{r}
my_dat <- trial_rm_sd %>% 
  add_column(block = NA) %>% mutate(block = as.factor(block))


my_dat$trial_name %>% unique()


results <- check_design_met(
  data = my_dat, #my_dat_1
  genotype = "accession_name",
  trial = "trial_name",
  traits = trait_ideal,
  rep = "rep_number",
  col = "col_number",
  row = "row_number",
  block = "block"
)

shared <- plot(results, type = "connectivity")

ggsave(paste('images\\shared_', trial_interest, Sys.Date(), ".png", sep = "_"),
       plot = shared, units = "in", dpi = 300, width = 8, height = 6)

summary <- results$summ_traits 

p1 <- summary %>% 
  ggplot(aes(x = traits , y = trial_name, label = round(miss_perc,2),  fill = miss_perc ))+
  geom_tile(color = "gray")+
  geom_text(color = "white")+
  theme_xiaofei() +
  labs(title = "Percentage of missing values (exp/trait)", x = "", y = "") 
  # theme(axis.text.x = element_text(hjust = 1 , angle = 75, size = 16),
  #       axis.text.y = element_text(size = 16))
p1
ggsave(paste("images\\missing_", trial_interest, Sys.Date(), ".png", sep = "_"),
     plot = p1, units = "in", dpi = 300, width = 15, height = 6
   )
master_data[["summ_traits"]] <- summary
```

## Single trial analysis

I should test removing the trials with non check clones. 

```{r}

obj <- single_trial_analysis(results = results,
                             progress = TRUE,
                             remove_outliers = FALSE)

#elements_to_remove <- c("202050DVPRG_ciat", "202136DVPRG_ciat")

trials <- unique(my_dat$trial_name)

header_sort = vector()
i = 1
for (i in 1:length(trials)) {
  
  cat("\n_______________")
  cat("\nTRIAL:", trials[i], "\n")
  cat("_______________\n")
  
  for (j in 1:length(trait_ideal)) {
    
    blue_blup <- obj$blues_blups %>% 
      filter(trial == trials[i]) %>% 
      select(-c(trial, seBLUEs, seBLUPs, wt)) %>% 
      pivot_wider(names_from = "trait", values_from = c("BLUEs", "BLUPs"))
    
    header_sort = c(header_sort,
                  grep(trait_ideal[j], sort(names(blue_blup)), value=TRUE))
  blue_blup <- blue_blup %>% dplyr::select(genotype, any_of(header_sort)) %>% 
    mutate(across(where(is.double), round, 1))
  }
 master_data[[paste0("BLUP_BLUE_", trials[i])]] <- blue_blup
}

# save into pdf spatial plots
if(TRUE){
  pdf(paste(here::here("output//"), "01_", trial_interest, "_spatial_plot_", 
            ".pdf", sep = ""), width = 8, height = 6)
  
  plot(obj, type = "spatial") 
  
  dev.off()
}
```

## Single heritability

```{r}
single_h2 <- obj$resum_fitted_model[ ,1:3] %>% 
  mutate(heritability = round(heritability, 2)) %>% 
  group_by(trial) %>%
  spread(trait, value = heritability) 

master_data[["single_h2"]] <- single_h2 # 202157DVGXE_polo H2 = 0.99

single_h2 %>% 
  write.table("clipboard", sep = "\t", col.names = T, row.names = F, na = "")


# single heritability plot
H2 <- obj$resum_fitted_model[ ,1:3] %>% 
  ggplot(aes(x = trait, y = trial, label = heritability, fill = heritability)) +
  geom_tile(color = "black") +  # This will also avoid drawing borders for NA tiles
  geom_text(color = "white") +  # Avoids placing text on NA tiles
  labs(title = "Broad sense heritability", x = "", y = "") +
  theme_xiaofei() +
  scale_x_discrete(position = "top") +
  theme(axis.text.x = element_text(angle = 75, hjust = 0, size = 10),
        axis.text.y = element_text(size = 13)) 
H2
ggsave(paste0("images\\H2", trial_interest, "_", ".png"),
     plot = H2, units = "in", dpi = 300, width = 15, height = 6
   )

# my_dat %>% filter(trial_name == "2023128DMEPR_bant") %>% 
#   select(rep_number, germinated_number_plot)

```

## Multi environmetal analysis

```{r}
# if (length(unique(my_dat$trial_name)) <= 3) { 
# traits_to_remove <- single_h2 %>% 
#   pivot_longer(!trial, names_to = "traits", values_to = "heritability") %>% 
#   filter(heritability < 0.1 | is.na(heritability)) %>% distinct(traits) %>% pull(traits)
# 
# # traits removed because of low heritability
# single_h2 %>% 
#   pivot_longer(!trial, names_to = "traits", values_to = "heritability") %>% 
#   filter(heritability < 0.1 | is.na(heritability)) %>% 
#   write.table("clipboard", sep = "\t", col.names = T, row.names = F)
# 
# met_results <- met_analysis(obj, 
#                             #filter_traits = trait_ideal[!trait_ideal %in% c(traits_to_remove)],
#                             h2_filter = 0.09,
#                             progress = TRUE)
# }

met_results <- met_analysis(obj,
                            #filter_traits = trait_ideal[!trait_ideal %in% c(traits_to_remove)],
                            h2_filter = 0.09,
                            progress = TRUE)


# h2 gxe
master_data[["h2_gxe"]] <- 
  met_results$heritability %>% 
  arrange(desc(h2)) %>%
  mutate(across(where(is.numeric), round, 2))


# BLUPs gxe
BLUPs_table <- 
  met_results$overall_BLUPs %>% 
  select(-c(std.error, status)) %>% 
  group_by(genotype) %>% 
  spread(trait, value = predicted.value) %>% 
  rename("accession_name" = genotype) %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  ungroup() 
#save the BLUPs data
master_data[[paste0("BLUPs_", "gxe")]] <- BLUPs_table
```

## Genotypic correlation

```{r}

geno_cor <- gg_cor(
  colours = c("red", "white", "blue"),
  data = BLUPs_table, # remove carotenoids
  label_size = 4
) + 
  theme(
  axis.text.y = element_text(size = 14),
  axis.text.x = element_text(size = 14))


ggsave(paste("images\\geno_corr", trial_interest, Sys.Date(), ".png", sep = "_"),
       units = "in", dpi = 300, width = 12, height = 8)
```

## Genotypic Correlation: Locations

```{r}
# Define a function to create plots and save them
plot_and_save <- function(trait, file_name, met_results, trial_interest) {
  covcor_heat(matrix = met_results$VCOV[[trait]]$CORR, size = 4, legend = c(0.35, 0.8)) +
    theme(axis.text.y = element_text(size = 14),
          axis.text.x = element_text(size = 14))
  
  # Construct file path using the file_name and trial_interest
  file_path <- paste0("images\\", file_name, "_vcor", trial_interest, Sys.Date(), ".png")
  
  # Save the plot
  ggsave(file_path, units = "in", dpi = 300, width = 8, height = 6)
}

# List of traits to plot
traits <- colnames(BLUPs_table)[!colnames(BLUPs_table) %in% c("accession_name")]

# Iterate over each trait and apply the function
for (trait in traits) {
  plot_and_save(trait, trait, met_results, trial_interest)
}
```

## Anova model

```{r, eval=FALSE, include=FALSE}
library(agricolae)
library(broom)

# List of variables for ANOVA

# removed variables presented in a single trial
variables <- trait_ideal #[!trait_ideal %in% c("lodging1_3_4mon", "lodging1_3_6mon", "plant_type_4mon")]

# factor 
my_dat$rep_number = as.factor(my_dat$rep_number)
my_dat$trial_name = as.factor(my_dat$trial_name)
my_dat$accession_name = as.factor(my_dat$accession_name)
my_dat$year = as.factor(my_dat$year)

my_dat <- my_dat %>%
  mutate(trial_name = as.character(trial_name),
         short_loc = substr(trial_name, nchar(trial_name) - 3, nchar(trial_name)))

my_dat$short_loc = as.factor(my_dat$short_loc)

# Run ANOVA for each variable using a for loop
for (variable in variables) {
  
  eqt_fixed <- reformulate(c("accession_name", 
                             "year",
                             "short_loc",
                             "rep_number:short_loc:year",
                             "short_loc:year",
                             "accession_name:year",
                             "accession_name:short_loc"
                             ), response = variable)
  
  # Run ANOVA and 
  anova_result <- aov(eqt_fixed, data = my_dat)
  
  
  

  out <- HSD.test(anova_result,"accession_name")

  out_year <- HSD.test(anova_result,"year")

  #out_loc <- HSD.test(anova_result,"short_loc")
  
  out_2WI <- HSD.test(anova_result, c("short_loc", "year"))
  
  #out_3WI <- HSD.test(anova_result, c("rep_number", "short_loc", "year"))
  
  # Store the result in the list
  # anova tables
  master_data[[paste("aov", variable, sep = "_") ]] <- 
    tidy(anova_result) #%>% 
    #mutate(across(where(is.double), round, 3)
           #)

  #HSD clones
  master_data[[paste("HSD_geno", variable, sep = "_") ]] <- 
    out$groups %>% rownames_to_column("accession_name") %>% 
    pivot_longer(variable, names_to = "trait", values_to = "mean") %>% 
    mutate(across(where(is.double), round, 3))
  
  # HSD year
  master_data[[paste("HSD_year", variable, sep = "_") ]] <- 
    out_year$groups %>% rownames_to_column("year") %>% 
    pivot_longer(variable, names_to = "trait", values_to = "mean") %>% 
    mutate(across(where(is.double), round, 3))
  
    # HSD loc
  # master_data[[paste("HSD_loc", variable, sep = "_") ]] <- 
  #   out_loc$groups %>% rownames_to_column("loc") %>% 
  #   pivot_longer(variable, names_to = "trait", values_to = "mean") %>% 
  #   mutate(across(where(is.double), round, 3))
  # 
  
    # HSD 2WI
  master_data[[paste("HSD_2WI", variable, sep = "_") ]] <- 
    out_2WI$groups %>% rownames_to_column("short_loc:year") %>% 
    pivot_longer(variable, names_to = "trait", values_to = "mean") %>% 
    mutate(across(where(is.double), round, 3))
  
      # HSD 3WI
  # master_data[[paste("HSD_3WI", variable, sep = "_") ]] <- 
  #   out_3WI$groups %>% rownames_to_column("rep_number:short_loc:year") %>% 
  #   pivot_longer(variable, names_to = "trait", values_to = "mean") %>% 
  #   mutate(across(where(is.double), round, 3))
} 
```

# extracting patterns of each result 

```{r}
# ANOVA
ind <- grep("^aov", names(master_data))

# accession_name HSD
ind_HSD <- grep("^HSD_geno", names(master_data))

# trial_name HSD
loca_HSD <- grep("^HSD_2WI", names(master_data))
```

# Plotting all accession means

```{r message=FALSE, warning=FALSE}
library(data.table)
mean_geno <- master_data[ind_HSD] %>% rbindlist()

mean_geno %>% ggplot(aes(x = fct_reorder(accession_name, trait), 
             y = mean)) +
  facet_wrap(~trait, scales = "free_y") +
  geom_col(col = "black") +
  scale_fill_jco() +
  geom_text(aes(label = groups), vjust = 1, col = 'white', size = 2) +
  labs(x = NULL, y = " ") +
  theme_xiaofei() +
   theme(
    axis.text.x = element_text(size = 8, vjust = 1, angle = 75),
    axis.text.y = element_text(size = 8),
    plot.title = element_text(color = "black"),
    strip.text.x = element_text(
      size = 7, face = "bold.italic"))

ggsave(paste("images\\geno_mean", trial_interest, ".png", sep = "_"),
     units = "in", dpi = 300, width = 18, height = 10
   )

```

# ploting accession means (single plot)

```{r}
for(i in 1: length(master_data[ind_HSD])){
  
  # Select current dataframe
  current_df <- master_data[ind_HSD][[i]]
  
  # Set limits
  y_max = max(current_df$mean, na.rm = TRUE) * 1.2
  y_min = 0
  
  # Built barplot
    plot_box = ggplot(current_df, aes(x = fct_reorder(accession_name, -mean), 
                                      y = mean))+
      geom_col(col = "black") +
      coord_cartesian(ylim = c(y_min, y_max))+
      scale_fill_jco() + 
      geom_text(aes(label = groups), vjust = -0.3, size = 2.2) + 
      labs(x = NULL, y = unique(current_df$trait)) + 
      theme_xiaofei() +
      theme(axis.text.x = element_text(vjust = 1)) 
    
    # Print out barplots
    plot(plot_box) 
    
#    Save barplots
    ggsave(paste("images/", unique(current_df$trait),  trial_interest, ".png", sep = "_")  , plot_box, width = 7, height = 4)
}
```

# Plotting all location means

```{r}

mean_loca <- master_data[loca_HSD] %>% rbindlist()

mean_loca %>% ggplot(aes(x = fct_reorder(`short_loc:year`, trait), 
             y = mean)) +
  facet_wrap(~trait, scales = "free_y") +
  geom_col(col = "black") +
  scale_fill_jco() +
  geom_text(aes(label = groups), vjust = 1, col = 'white') +
  labs(x = NULL, y = " ") +
  theme_xiaofei() +
   theme(
    axis.text.x = element_text(size = 8, vjust = 1, angle = 75),
    axis.text.y = element_text(size = 8),
    plot.title = element_text(color = "black"),
    strip.text.x = element_text(
      size = 7, face = "bold.italic"))

ggsave(paste("images\\mean_loca", trial_interest, ".png", sep = "_"),
     units = "in", dpi = 300, width = 15, height = 10
   )

```

## Save the master data results

```{r}
folder_output <- here::here("output//")
meta_file_name <- paste0(folder_output, paste("2023", trial_interest, "master_results", Sys.Date(), ".xlsx", sep = "_"))

write.xlsx(master_data, file = meta_file_name)
```

# AMMI analysis

```{r}
library(metan)

# create year_loc var

my_dat <- my_dat %>% mutate(year_loc = paste(short_loc, year, sep = "_"))

variables_ideal <- c("yield_ha", "starch_content", "starch_yield_ha")
  
  # variables[!variables %in% c("CMD_infect_plants", "CMD_incidence"
  #                                              #"DM_nirs", "DM_raw", "frogskin1_5"
  #                                              )]

ammi_results <- list()

for (i in 1:length(variables_ideal)) {
  current_variable <- variables_ideal[i]
  
  # Make sure to pass the variable dynamically to the AMMI model function
  ammi_model <- performs_ammi(my_dat, year_loc, accession_name, rep_number, !!rlang::sym(current_variable))
  
  # Use double bracket indexing for dynamic variable names
  ammi_results[[paste0("AMMI_", current_variable)]] <- ammi_model[[current_variable]]$ANOVA
  
  # Plot the scores and add the custom theme
  a <- plot_scores(ammi_model, size.text.env = 1, size.tex.gen = 2) + theme_xiaofei()
  a_2 <- plot_scores(ammi_model, type = 2, polygon = TRUE, size.text.env = 1,
                     size.tex.gen = 2) + theme_xiaofei()
  
  # Fix the typo in 'size.tex.env', it should be 'size.text.env', and use TRUE instead of T
  a_3 <- plot_scores(ammi_model, type = 4, repulsion = 2, repel = TRUE,
                     size.text.env = 3, size.tex.gen = 1.5, 
                     max_overlaps = 25, x.lab = "Environment PC1", 
                     y.lab = current_variable, title = FALSE) + theme_xiaofei() +
            theme(legend.position = "none") 
  
  # Save the plots with corrected file path construction
  ggsave(paste0("images/AMMI2_", current_variable, "_", trial_interest, ".png"), 
         plot = a_2, width = 6, height = 4)
  
  ggsave(paste0("images/AMMI1_", current_variable, "_", trial_interest, ".png"), 
         plot = a, width = 6, height = 4)
  
  ggsave(paste0("images/stability_", current_variable, trial_interest, ".png"), 
         plot = a_3, width = 6, height = 4)
}
```


## Save the master data AMMI results

```{r}
meta_file_name <- paste0(folder_output, paste("2023", trial_interest, "ammi_results", Sys.Date(), ".xlsx", sep = "_"))

write.xlsx(ammi_results, file = meta_file_name)
```

# Locations map

```{r}
library(ggrepel)
library(maps)


# Get map data for the relevant countries

world <- map_data("world")

countries_data <- world[world$region %in% c("Cambodia", "Laos", "Vietnam", "Thailand"), ]


cambodia_map <- map_data("world", region = "Cambodia")
laos_map <- map_data("world", region = "Laos")
vietnam_map <- map_data("world", region = "Vietnam")
thailand_map <- map_data("world", region = "Thailand")

# Create a data frame with your locations and their approximate coordinates
# Create a data frame with your locations and their approximate coordinates using short names
locations <- data.frame(
  Location = c("cham", "bant", "rata", "anlo", "stun", "naph", "blik", "that", "laog"),
  Latitude = c(12.5, 13.8, 13.1, 14.1, 13.5, 18.0, 18.2, 15.2, 14.6),
  Longitude = c(105.3, 103.0, 103.2, 104.1, 106.1, 102.0, 104.5, 106.5, 106.3)
)


# Combine the map data for a single plot
southeast_asia_map <- rbind(cambodia_map, laos_map, vietnam_map, thailand_map)

# Create a ggplot object for the map
static_map <- ggplot() +
  # Plot the countries
  geom_polygon(data = countries_data, aes(x = long, y = lat, group = group),
               fill = "lightblue", color = "black", linewidth = 0.03)  +
  # Plot the locations in Malawi
  geom_point(data = locations, aes(x = Longitude, y = Latitude), size = 2, color = "red") +
  geom_text_repel(data = locations, aes(x = Longitude, y = Latitude, label = Location),
                  vjust = -0.5, size = 3) +
  # Add labels for the countries
  geom_text(data = countries_data %>% filter(region == "Thailand"), aes(x = mean(long), 
                                                                        y = 15, label = region), 
            size = 3, fontface = "bold", color = "darkblue", stat = "unique") +
  # Add labels for the countries
  geom_text(data = countries_data %>% filter(region == "Cambodia"), aes(x = 103.8, 
                                                                        y = 11.5, label = region), 
            size = 3, fontface = "bold", color = "darkblue", stat = "unique") +
  labs(x =  "Latitude", y = "Longitude") +
    # Add labels for the countries
  geom_text(data = countries_data %>% filter(region == "Laos"), aes(x = 102.8, 
                                                                        y = 19, label = region), 
            size = 3, fontface = "bold", color = "darkblue", stat = "unique") +
  # Add labels for the countries
  geom_text(data = countries_data %>% filter(region == "Vietnam"), aes(x = 105, 
                                                                        y = 21.5, label = region),
            size = 3, fontface = "bold", color = "darkblue", stat = "unique") +
  labs(y =  "Latitude", x = "Longitude") +
  labs(title = "Locations in Southeast Asia", subtitle = "Various districts and provinces") +
  theme(plot.title = element_text(face = "bold.italic"),
        plot.subtitle = element_text(face = "italic")) +
  coord_fixed(1.3) # This sets the aspect ratio

# Display the map
print(static_map)

ggsave(paste("images\\map", trial_interest, ".png", sep = "_"),
       plot = static_map, units = "in", dpi = 300, width = 6, height = 5
)
```




