---
title: "Get tidy data"
author: "Luis Fernando Delgado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("https://raw.githubusercontent.com/Cassava2050/PPD/main/utilities_tidy.R")
```

```{r}
# Define the function to read and load multiple CSV files from cassavabase
read_cassavabase <- function(phenotypeFiles, skip_col) {
  possibly_read_csv <- purrr::possibly(read.csv, NULL)
  
  # Initialize a list to store data from each file
  all_data <- list()
  
  # Loop through each file and read the data
  for (file in phenotypeFiles) {
    indata <- possibly_read_csv(file, 
                                na.strings = c("#VALUE!", NA, ".", "", " ", "-", "\""), 
                                stringsAsFactors = FALSE)
    
    # If the initial read fails (returns NULL), try again with the skip_col parameter
    if (is.null(indata)) {
      indata <- possibly_read_csv(file, 
                                  na.strings = c("#VALUE!", NA, ".", "", " ", "-", "\""), 
                                  stringsAsFactors = FALSE, 
                                  skip = skip_col, 
                                  check.names = FALSE)
    }
    
    # Check if the data was successfully read
    if (!is.null(indata)) {
      cat("\nTrials interested are:\n", unique(indata[, 6]))
      # Append the data to the list
      all_data[[file]] <- indata
    } else {
      cat("\nFailed to read the file: ", file)
    }
  }
  
  # Identify the union of all column names across all data frames
  all_columns <- unique(unlist(lapply(all_data, names)))
  
  # Add missing columns to each data frame with NA values
  all_data <- lapply(all_data, function(df) {
    missing_cols <- setdiff(all_columns, names(df))
    df[missing_cols] <- NA
    df <- df[all_columns]  # Reorder columns to ensure they match
    return(df)
  })
  
  return(all_data)
}

local_file <- "yes"
# Define the parameters for the files
if (local_file == "yes") {
  folder <- here::here("data//")
  files <- c("phenotype.csv", 
             "phenotype(45).csv", 
             "phenotype(46).csv",
             "phenotype(47).csv",
             "phenotype(48).csv",
             "phenotype(49).csv",
             "phenotype(50).csv",
             "phenotype(51).csv",
             "phenotype(52).csv",
             "phenotype(53).csv")
  skip_col <- 3 # double check the number of col skipped
  trial_interest <- "DMEPR"
  year_interest <- 2023
}

# Construct the full file paths
file_paths <- paste0(folder, files)

# Load the data from multiple files
sel_data <- read_cassavabase(phenotypeFiles = file_paths, skip_col = skip_col)

# If you want to combine all data frames into one, you can do this:
combined_data <- do.call(rbind, sel_data)
row.names(combined_data) <- NULL
```

# ---- Change columns into standar names ----

```{r}
sel_data_kp <- change_colname(combined_data, NA)
```


## change the column class

```{r}
obs_col <- c(
  names(sel_data_kp)[str_detect(names(sel_data_kp), "obs_")],
  "use_rep_number", "blockNumber",
  "use_plot_number"
)
sel_data_kp %<>%
  mutate(across(all_of(obs_col), as.numeric))

# remove - , replace by _
names(sel_data_kp) = gsub("-", "_", names(sel_data_kp))
```

## Duplications in row and cols

```{r}
duplicated_plot <- row_col_dup(sel_data_kp)
```

## Plot trial layout
```{r}
trial_layout <- function(trial = sel_data_kp) {
  trial_list <- unique(trial$use_trial_name)
  for (i in 1:length(trial_list)) {
    trial_i <- trial %>%
      filter(use_trial_name %in% trial_list[i])
    myplot <- ggplot(trial_i, aes(x = factor(use_col_number), y = factor(use_row_number), fill = factor(use_rep_number))) +
      geom_tile(color = "black", linewidth = 0.5) + # Black border on tiles
      # geom_tile(
      #   data = trial_i %>% filter(use_check_test == "check"),
      #   aes(fill = use_check_test), col = "black"
      # ) +
      scale_fill_jco() +
      geom_text(
        data = trial_i %>% filter(use_check_test == "check"),
        aes(label = use_accession_name), size = 2
      ) +
      geom_text(
        data = trial_i %>% filter(use_check_test == "test"),
        aes(label = use_plot_number), size = 3
      ) +
      labs(x = "col_number", y = "row_number", fill = "rep", title = trial_list[i]) +
      coord_fixed() + # Square tiles
      theme_xiaofei()

    print(myplot)
    
    ggsave(paste("images\\layout", trial_list[i], ".png", sep = "_"),
       plot = myplot, units = "in", dpi = 300, width = 14, height = 12)
  }
}
trial_layout(sel_data_kp)
```

## Check the clone name
```{r}
cloneName_new_old <- check_clone_name(
  clone_list = sel_data_kp$use_accession_name,
  new_names = NA,
  add_check = NULL
)

trial_standard <- sel_data_kp %>%
  left_join(cloneName_new_old,
            by = c("use_accession_name" = "accession_name_ori")
  ) %>%
  select(-use_accession_name) %>%
  rename(use_accession_name = use_accession_name.y)
```

## Add GIS data

```{r}
trial_standard <- add_GIS(trial_standard)
```

# extract checks

```{r}
accession_rep_ct <- trial_standard %>%
  count(use_trial_name, use_accession_name, use_rep_number)  %>%
  arrange(use_trial_name) %>%
  filter(n>1)
accession_rep_ct 


conducted_trials <- 
  trial_standard %>% group_by(use_trial_name, use_plant_date,use_harvest_date, use_location) %>% 
  summarise(n_gen = n_distinct(use_accession_name)) %>% 
  mutate(harvesting_time = 
           interval(ymd(use_plant_date), ymd(use_harvest_date)) %>% as.period,
         harvesting_time = paste0(harvesting_time@month, "month ", harvesting_time@day, "day")) %>% 
  ungroup()

conducted_trials

conducted_trials %>% relocate(harvesting_time, .after = use_harvest_date) %>% 
  write.table("clipboard", sep="\t", col.names = T, row.names = F)
```


## plot plant number

```{r}
plants_plot <- trial_standard %>%
  group_by(use_trial_name) %>%
  count(obs_planted_number_plot) 
plants_plot

```


## Frequency harvest plant number

```{r}
plants_harvested <- trial_standard %>%
  group_by(use_trial_name) %>%
  count(obs_harvest_number) %>% arrange(desc(obs_harvest_number))
```


# Planted and harvested

```{r}
plants_plot %>% select(-n) %>% 
  left_join(plants_harvested %>% 
              summarise(harvested_plants = max(obs_harvest_number, na.rm = TRUE)), 
            by = "use_trial_name") %>% 
  write.table("clipboard", sep="\t", col.names = T, row.names = F)
```

# Harvested plants

```{r}
plants_to_harvest <- plants_harvested %>% 
  ggplot(aes(x = factor(obs_harvest_number), 
             y = n, fill = factor(obs_harvest_number))) +
  geom_col(col = 'black') +
  theme_xiaofei() +
  theme(legend.position="top") +
  theme(
    axis.text.x = element_text(size = 5, vjust = 1, angle = 65))+
  labs(x = "Harvest_plant_number", y = "Freq", fill = "Harvest_plant_number") +
  facet_wrap(~ use_trial_name, ncol = 5)

ggsave(paste("images\\bar", trial_interest, ".png", sep = "_"),
       plot = plants_to_harvest, units = "in", dpi = 300, width = 9, height = 6)
```

## Compute germination and yield


```{r}
trial_standard_new <- trial_standard %>%
  mutate(
    obs_germination_perc =
      obs_germinated_number_plot / obs_planted_number_plot * 100,
    obs_starch_yield_ha =
      obs_starch_content * obs_yield_ha / 100
  )
```

## Is numeric all traits?
Checks if all traits in the dataset are numeric, ensuring data type consistency for analysis.


```{r}
is_numeric(trial_data = trial_standard_new)
```

## Get the tidy data
Prepares the final tidy dataset for analysis by selecting and renaming variables, exemplifying data tidying practices.

```{r}
meta_info = names(trial_standard_new)[str_detect(names(trial_standard_new), "use_")]
meta_info = gsub("use_", "", meta_info)
meta_info
trial_tidy = trial_standard_new
names(trial_tidy)= gsub("use_", "", names(trial_standard_new))
# observations
trait_list = names(trial_tidy)[str_detect(names(trial_tidy), "obs_")]
trait_list = gsub("obs_", "", trait_list)
trait_list
names(trial_tidy)= gsub("obs_", "", names(trial_tidy))
trial_tidy = trial_tidy[c(meta_info, trait_list)]
```

# Boxplots and grouping boxplots
Generates boxplots for trait distributions across trials.

```{r warning=FALSE}
trait_wanted <- trait_list
boxplot_traits(my_dat = trial_tidy, 
               folder = here::here("output//"), 
               trait_wanted = trait_wanted, 
               trial_interest = trial_interest)

trial_tidy %>% filter(CMD_harvest > 5) %>% 
  select(trial_name, location, accession_name, plot_name, CMD_harvest) %>% 
  write.table("clipboard", col.names = T, row.names = F, sep = "\t")
```

##Deleting the CMD outlier

```{r}
trial_tidy <- trial_tidy %>% filter(!CMD_harvest > 5)
```

## Grouping boxplot

```{r}

plot_bxp <- trial_tidy %>%
  pivot_longer(
    cols = all_of(trait_wanted),
    names_to = "var",
    values_to = "values"
  ) %>%
  filter(!var %in% c(
    "stake_plant", "planted_number_plot",
    "harvest_number", "root_weight_air",
    "root_weight_water", "harvest_number_plan",
    "yield_ha_v2", "root_rot_perc", "flesh_color_50inner1_8",
    "lodging1_3_6mon", "lodging1_3_4mon", "root_rot_number",
    "shoot_weight_plot"
  )) %>%
  ggplot(aes(x = trial_name, y = values)) +
  geom_violin(fill = "gray") +
  geom_boxplot(width = 0.2) +
  labs(x = NULL, y = NULL, title = "") +
  theme_xiaofei() +
  theme(
    axis.text.x = element_text(size = 8, vjust = 1, angle = 75),
    axis.text.y = element_text(size = 8),
    plot.title = element_text(color = "black"),
    strip.text.x = element_text(
      size = 7, face = "bold.italic")
  ) +
  facet_wrap(~var,
    ncol = 3, scales = "free_y"
  ) 
print(plot_bxp)

ggsave(paste0("images\\boxplot_fixed_", trial_interest, Sys.Date(), ".png"),
    plot = plot_bxp, units = "in", dpi = 300, width = 14, height = 10
  )

```

## Save the tidy data for analysis.
Write the cleaned and processed dataset to a CSV file for further analysis.

```{r eval=FALSE}
write.csv(trial_tidy, here::here("output", paste("01_", year_interest, trial_interest,
  "_tidy_data4analysis_", Sys.Date(), ".csv", sep = "")), row.names = FALSE)
```



